Multitask matrix completion for learning protein interactions across diseases

### 2016-11-22 paper.tex
new stuff

\author{Meghana Kshirsagar$^{1}$, Keerthiram Murugesan$^{2}$, Jaime G. Carbonell$^{2}$ and Judith Klein-Seetharaman$^{3}$}
%\address{$^{1}$Memorial Sloan Kettering Cancer Center, 1275 York Ave., NY 10021\\
%$^{2}$Language Technologies Institute, Carnegie Mellon Univ., 5000 Forbes Ave., Pittsburgh PA 15213, USA\\
%$^{3}$Metabolic \& Vascular Health, Warwick Medical School, Univ. of Warwick, Coventry, UK
%}

\begin{abstract}
\textbf{Motivation:}
Disease causing pathogens such as viruses, introduce their proteins into the host cells 
where they interact with the host's proteins enabling the virus to replicate inside the host.
These interactions between pathogen and host proteins are key to understanding infectious diseases.
Often multiple diseases involve phylogenetically related or biologically similar pathogens. Here we present a multitask learning method to jointly model interactions between human proteins and three different, but related viruses: \textit{Hepatitis C}, \textit{Ebola} virus and \textit{Influenza A}. \\
\textbf{Results:}
Our multitask matrix completion based model uses a shared low-rank structure in addition to a task-specific sparse structure to incorporate the various interactions.
We obtain between 7 to 39 percentage points improvement in predictive performance over prior state-of-the-art models. 
We show how our model's parameters can be interpreted to reveal both general and specific
interaction-relevant characteristics of the viruses. \\
\textbf{Availability:}
Code and data is available at: \url{http://www.cs.cmu.edu/~mkshirsa/bsl_mtl.tgz}\\
\textbf{Contact:}
\href{mkshirsa@cs.cmu.edu}{meghana.ksagar@gmail.com}
\end{abstract}

@@

\begin{figure}
\includegraphics[scale=0.25, trim = 1.1cm 0.9cm 0 0]{pca_bslmtl_with_clusters.eps}
\caption{Principal component analysis (PCA) of virus proteins in the projected subspace. The first two principal components are shown. Shape of the points indicates which virus that protein comes from. We have highlighted three clusters for which we show GO term enrichment analysis results in Table \ref{goterms}.}%
\label{fig:pca2}
\end{figure}

@@

\bibliographystyle{natbib}
\begin{normalsize}
\bibliography{references}
\end{normalsize}

>>>>

\author{Meghana Kshirsagar$^{1*}$, Keerthiram Murugesan$^{2}$,\\  Jaime G. Carbonell$^{2}$, Judith Klein-Seetharaman$^{3}$
\thanks{$^{1}$Memorial Sloan Kettering Cancer Center, 1275 York Ave., NY 10021}
\thanks{$^{2}$Language Technologies Institute, Carnegie Mellon Univ., 5000 Forbes Ave., Pittsburgh PA 15213, USA}
\thanks{$^{3}$Metabolic \& Vascular Health, Warwick Medical School, Univ. of Warwick, Coventry, UK}
}

\begin{abstract}
\textbf{Motivation:}
Disease causing pathogens such as viruses, introduce their proteins into the host cells 
where they interact with the host's proteins enabling the virus to replicate inside the host.
These interactions between pathogen and host proteins are key to understanding infectious diseases.
Often multiple diseases involve phylogenetically related or biologically similar pathogens. Here we present a multitask learning method to jointly model interactions between human proteins and three different, but related viruses: \textit{Hepatitis C}, \textit{Ebola} virus and \textit{Influenza A}. \\
\textbf{Results:}
Our multitask matrix completion based model uses a shared low-rank structure in addition to a task-specific sparse structure to incorporate the various interactions.
We obtain between 7 to 39 percentage points improvement in predictive performance over prior state-of-the-art models. 
We show how our model's parameters can be interpreted to reveal both general and specific
interaction-relevant characteristics of the viruses. \\
\textbf{Availability:}
Code and data is available at: \url{http://www.cs.cmu.edu/~mkshirsa/bsl_mtl.tgz}\\
\textbf{Contact:}
\href{mkshirsa@cs.cmu.edu}{kshirsam@mskcc.org}
\end{abstract}

@@

\begin{figure}
\includegraphics[scale=0.25, trim = 1.1cm 0.9cm 0 0]{Kshirsagar.Fig_2.eps}
\caption{Principal component analysis (PCA) of virus proteins in the projected subspace. The first two principal components are shown. Shape of the points indicates which virus that protein comes from. We have highlighted three clusters for which we show GO term enrichment analysis results in Table \ref{goterms}.}%
\label{fig:pca2}
\end{figure}

@@

\bibliographystyle{natbib}
\bibliography{references}
